### 2019-11-01 Paper_shortened_pdf.Rmd
Add DSEM results and update action letter

A potential confound might be introduced when operationalizing quantitative involvement in terms of both parents' absolute involvement: Effects that appears to be due to fathers' greater involvement might be due to children receiving **more** total care [see @deutschPaternalParticipationChild2001]. Examining the childcare distribution instead, and controlling for total care, mitigates this confound. Some prior research has attempted to measure the childcare distribution directly, by asking which parent is most responsible for certain tasks [@krollEarlyFatherInvolvement2016]. When quantitative involvement has been measured in terms of absolute involvement, latent means and differences can be used to statistically disentangle the childcare distribution from total joint care [@mcardleLatentVariableModeling2009].

@@

The amount of time parents spend with children is associated with the quality of parenting behaviors  [@sabattiniRelationMothersFathers2004]. Indeed, more involved parents tend to experience greater parenting self-efficacy, possibly reflecting the effects of experience [@gialloMothersFathersInvolvement2013a]. Furthermore, when both parents are more involved, they monitor each other's parenting behaviors to ensure that children are raised well [@mclanahanConsequencesFatherAbsence2014]. Consequently, we might expect greater total joint care to positively predict both parents' parenting quality.

@@

Sex differences are a relevant factor in the interplay between parenting and emotional adjustment. Parents tend to raise boys and girls differently; expressing more control with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and more negative parenting with boys [@strausCorporalPunishmentAmerican1999]. Furthermore, social learning theory posits that the same-sex parent plays a stronger role in socialization; particularly for sex-typed behavior, such as emotion display rules [@busseySocialCognitiveTheory1999]. Finally, mounting evidence suggests that parental sensitivity is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a]. We might thus expect parents to display more responsiveness, structure, and inductive reasoning with girls, and more negative parenting with boys. Furthermore, we might expect associations between qualitative parenting and emotional adjustment to be stronger between children and the same-sex parent.

Although sex differences in emotional adjustment are sometimes found in older children [@silkAdolescentsEmotionRegulation2003; @vanlissaRoleFathersMothers2018a], a meta-analysis found little evidence for such sex differences in young children [@else-questGenderDifferencesTemperament2006a]. We consequently did not expect to find sex differences in children's average levels of emotional adjustment.

@@

The present study aims to make several contributions. First, we include both mothers and fathers, in order to tease apart their common and unique roles in children's emotional adjustment development. Second, we incorporate both qualitative and quantitative measures of involvement. Third, we disentangle stable between-family differences from dynamic within-family predictive processes, in order to more accurately assess the direction of effects between qualitative involvement and emotional adjustment. Fourth, we examined interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work.

@@

To address our research questions, we used multilevel, autoregressive structural equation modeling [@asparouhovDynamicStructuralEquation2018]. In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean). This has several advantages: First, controlling for stable between-family differences allows us to estimate the direction of within-family effects more accurately [@hamakerCritiqueCrosslaggedPanel2015]. Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts. Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects. In this way, we included total joint childcare, the childcare distribution, and child sex as moderators. We conducted separate analyses for the four qualitative parenting dimensions, because a model including all four variables at once was too complex to estimate.

@@

tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
tabchildeff$Parameter[c(1,4)] <- c("Random intercepts of child effects on:", "Variance of child effects on:")
tab <- rbind(tabparenteff, tabchildeff)
#tab[is.na(tab)] <- ""
rownames(tab) <- NULL
reg_table(tab, caption = "Within-family predictive effects", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

We examined regression coefficients from total joint involvement and relative involvement to the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). With regard to parenting, we found partial support for our hypotheses: For both parents, joint involvement positively predicted greater warmth, consistency, and inductive reasoning. Notably, the association between total joint involvement and warmth was almost double in size for fathers, as compared to mothers. We did not find the hypothesized negative effects of joint total involvement on parental anger. Counter to our hypothesis, we found that greater total joint involvement was negatively associated with emotional adjustment. Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.

@@

```{r tabinvolvement, results = "asis"}
tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
tabinv$Parameter[c(1,5)] <- c("Effects of joint quantitative involvement on random intercepts of:",
                              "Effects of relative quantitative involvement on random intercepts of:")
tabinv$Parameter <- gsub("^(.+) on.*$", "\\1", tabinv$Parameter)
#reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
tabinvmod <- rbind(
  tab_std_sections$`Moderating effect of quantitative involvement on parenting effects`[c(1:3, 1, 4:5),],
  tab_std_sections$`Moderating effect of quantitative involvement on child effects`[c(1:3, 1, 4:5),])
tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on parenting effects of:",
                              "Moderating effects of relative involvement on parenting effects of:",
                              "Moderating effects of joint involvement on child effects on:",
                              "Moderating effects of relative involvement on child effects on:")
tab <- rbind(tabinv, tabinvmod)
#tab[is.na(tab)] <- ""
rownames(tab) <- NULL
reg_table(tab, caption = "Effects of quantitative involvement", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

To examine interactions between involvement quality and quantity, we tested whether between-family differences in quantitative involvement predicted the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth. Specifically, the within-families predictive effect of maternal warmth on emotional adjustment was negatively attenuated when both parents' total joint involvement was higher. This same effect was positively attenuated, however, if mothers' involvement was relative higher than fathers' involvement. We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.

With regard to the moderation of child effects by quantitative involvement, we found that the effect of adjustment on maternal warmth was negatively attenuated when parents' joint involvement was higher, and also when mothers' relative involvement was greater compared to fathers. We found no effects for consistency. For inductive reasoning, we found that that the cross-lagged effect of adjustment on maternal inductive reasoning was negatively attenuated when joint involvement was higher. Adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers. Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment more strongly predicted decreased paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning. We found no moderating effects for anger.

@@

```{r tabsex, results = "asis"}
tabsex <- rbind(
  tab_std_sections$`Sex differences in parenting and adjustment`,
  tab_std_sections$`Sex differences in cross-lagged effects`[c(1:3, 1, 4:5),])
tabsex$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabsex$Parameter)
tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
tabsex$Parameter[c(1,5,8)] <- c("Sex differences in parenting and adjustment",
                              "Sex differences in parenting effects of:",
                              "Sex differences in parenting effects on:")
#tabsex[is.na(tabsex)] <- ""
rownames(tabsex) <- NULL
reg_table(tabsex, caption = "Sex differences", note = "Sex is a binary predictor (0 = female, 1 = male). Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

Our operationalization of quantitative involvement, in terms of total joint care and the childcare distribution, allowed us to addresses the claim that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001]. However, counter to our hypothesis, joint care was *negatively* associated with between-family differences in emotional adjustment. One potential explanation is suggested by self-determination theory: Excessive levels of involvement might undermine children's autonomy needs, thereby leading to increased emotional maladjustment [@vandergiessenPerceivedAutonomySupport2014]. Research on "helicopter parenting" supports this interpretation, as psychological well-being was found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014]. Another potential explanation is that the amount of quantitative care provided is not only driven by parents' motivation and ability to be involved in their children's lives, but also by children's objective needs. Factors related to children's objective needs - such as developmental delay, or atypical development - might thus drive an increase in parents' quantitative involvement, and might also be associated with children's emotional maladjustment, thereby resulting in a negative correlation between the two.

@@

We found limited evidence that quantitative involvement moderated within-family parenting effects. One such effect is that the effect of maternal warmth on adjustment was contingent on both total joint care and the childcare distribution. Mothers' warmth predicted emotional adjustment less strongly when both parents were more involved, and more strongly when mothers were relatively more involved than fathers. These findings suggests a compensatory effect, whereby mothers' warmth can make up for disadvantages incurred by having a relatively uninvolved father [@shinnFatherAbsenceChildren1978].

The evidence that quantitative moderates child effects was somewhat stronger. Firstly, when total joint involvement was higher, and when mothers' relative involvement was greater compared to fathers, emotional adjustment became less predictive of maternal warmth. These findings can be interpreted in the context of the sizable average positive effect of adjustment on maternal warmth. In families where mothers' involvement was high - either in terms of joint, or relative involvement - this positive effect was attenuated somewhat. Secondly, total joint involvement positively attenuated the link from emotional adjustment to maternal inductive reasoning. This suggests that mothers have more of a role as disciplinarian in families where both parents are highly involved. When mothers were relatively more involved than fathers, fathers decreased their inductive reasoning more strongly in response to children's positive adjustment, and increased it more strongly in response to maladjustment. Together, these findings suggests that fathers take more of a role as disciplinarian when mothers are relatively more involved in care tasks.

@@

Historically, developmental theory has focused on how parents cultivate their children's development. Nowadays, it is commonly acknowledged that influences are reciprocal. However, commonly used developmental methods have been criticized for misrepresenting the direction of effects when stable between-family differences are not accounted for [@hamakerCritiqueCrosslaggedPanel2015]. One recent study with a similar design to the present study found that child effects substantially outnumbered parenting effects, and speculated that this might be due to the greater equality of parent-adolescent relationships, compared to earlier childhood [@vanlissaRoleFathersMothers2018a]. The present study casts doubt on this speculation, as we found that child effects outweighed parenting effects, even at a young age. The fact that both the present study and this previous work found a preponderance of child effects, despite the different age ranges of the samples, invites the question: Are parents really shaping children's development, or are they, instead, adapting to it? An extreme interpretation of our findings might be that parents "set the stage" for children's development, through their stable differences from other parents, and from there on, are merely responding to children's changing developmental needs. This extreme view can be nuanced by taking into consideration that not all causal effects are evident from the interplay between within-individual fluctuations. For example, although the genetic makeup that is passed from parents to children is stable, the expression of some genes emerges over time. Consequently, stable parent characteristics can cause change in children. Future research might examine alternative theoretical conceptualizations of parental influences in child development. Another point to consider is that parent effects on children might play out on a different time scale than was captured by these macro-longitudinal studies. If parents shape their children through specific behaviors in dyadic interactions, then we might fail to capture parenting effects when using annual reflective self-reports.

@@

Several strengths allow us to make unique contributions to the literature. First, mothers and fathers both completed identical parenting measures, which allowed us to disentangle and compare their unique roles in children's emotional adjustment development. Second, we incorporated both quantitative and qualitative involvement. This allowed us to provide a more nuanced overview of the links between parental involvement and emotional adjustment development than prior work, including how the distribution of quantitative childcare relates to parenting quality and emotional adjustment, and moderates links between these two constructs. Although some prior studies have also incorporated both quality and quantity of parenting, this work did not disentangle the total amount of care children received from the relative distribution of care tasks [see @deutschPaternalParticipationChild2001]. We addressed this limitation by re-expressing the variance of maternal and paternal quantitative involvement into joint total involvement, and differential involvement. Third, we used innovative methods to examine the extent to which parenting and adjustment are linked at the between- versus within-family level. Compared to methods that do not distinguish between these levels, our approach estimates dynamic reciprocal processes within the family more accurately. Our findings contribute to a growing literature suggesting that stable between-family differences might play a larger role than was previously thought, even in early childhood. Confidence in the results is bolstered by the fact that we used multi-informant reports: Each parent reported on their own parenting, and teachers reported on children's emotional adjustment. This avoids common method bias and single reporter bias. As teachers are likely to be "blind" to parents' child-rearing behavior, and have no stake in the parent-child relationship, their reports might be less biased in general.

@@

@@

We found substantial evidence that both parents reported engaging more in behaviors that are positively associated with children's adjustment when total joint involvement was higher. Conversely, when mothers were more involved than fathers, mothers' expressed more warmth, but fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth). Children's emotional adjustment, too, was lower when mothers were more involved than fathers. These findings highlight the important of considering interactions between quantitative and qualitative involvement, and suggest that all family members might benefit from a more equal distribution of childcare tasks.

>>>>

A potential confound might be introduced when operationalizing quantitative involvement in terms of both parents' absolute involvement: Effects that appears to be due to fathers' greater involvement might be due to children receiving **more** total care [see @deutschPaternalParticipationChild2001]. This confound can be mitigated by examining the effect of the distribution of childcare, while controlling for total care. Some prior research has attempted to achieve this by assessing the childcare distribution directly, by asking which parent is most responsible for certain tasks [@krollEarlyFatherInvolvement2016; @keizerInfluenceFathersMothers2019]. Alternatively, when quantitative involvement has been measured in terms of absolute involvement, one can estimate a latent mean and difference to statistically disentangle joint involvement from relative involvement [@mcardleLatentVariableModeling2009]. Joint involvement reflects the total amount of care provided to children, and relative involvement reflects the childcare distribution.

@@

The amount of time parents spend with children is associated with the quality of parenting behaviors  [@sabattiniRelationMothersFathers2004]. Indeed, more involved parents tend to experience greater parenting self-efficacy, possibly reflecting the effects of experience [@gialloMothersFathersInvolvement2013a]. Furthermore, when both parents are more involved, they monitor each other's parenting behaviors to ensure that children are raised well [@mclanahanConsequencesFatherAbsence2014]. Consequently, we might expect greater joint involvement to positively predict both parents' parenting quality.

@@

Sex differences are a relevant factor in the interplay between parenting and emotional adjustment. Parents tend to raise boys and girls differently; expressing more control with girls [@pomerantzRoleMaternalControl1998; @vanlissaRoleFathersMothers2018a], and more negative parenting with boys [@strausCorporalPunishmentAmerican1999]. Furthermore, social learning theory posits that the same-sex parent plays a stronger role in socialization; particularly for sex-typed behavior, such as emotion display rules [@busseySocialCognitiveTheory1999]. Finally, mounting evidence suggests that parental sensitivity is greatest in mother-daughter relationships, and lowest in father-son relationships [see @vanlissaRoleFathersMothers2018a]. We therefore hypothesize that parents will display more responsiveness, structure, and inductive reasoning towards girls, and more negative parenting towards boys. Furthermore, we predict stronger associations between qualitative parenting and emotional adjustment between children and the same-sex parent.

Although sex differences in emotional adjustment are sometimes found in older children [@silkAdolescentsEmotionRegulation2003; @vanlissaRoleFathersMothers2018a], a meta-analysis found little evidence for such sex differences in young children [@else-questGenderDifferencesTemperament2006a]. Consequently, we have no hypothesis regarding sex differences in children's average levels of emotional adjustment.

@@

The present study aims to make several contributions. First, we include both mothers and fathers, in order to tease apart their common and unique roles in children's emotional adjustment development. Second, we incorporate both qualitative and quantitative measures of involvement. Third, we disentangle stable between-family differences from dynamic within-family predictive processes, in order to more accurately assess the direction of effects between qualitative involvement and emotional adjustment. Fourth, we examined interactions between qualitative and quantitative involvement, which have been largely overlooked in prior work. Finally, we examined sex as a moderator of between-family differences in all study variables and within-family predictive effects.

@@

To address our research questions, we used multilevel, autoregressive structural equation modeling [@asparouhovDynamicStructuralEquation2018]. In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean). This has several advantages: First, controlling for stable between-family differences allows us to estimate the direction of within-family effects more accurately [@hamakerCritiqueCrosslaggedPanel2015]. Any stable factors which might bias cross-lagged effects - such as the influence of SES or neighborhood effects - are accounted for at the between level, and reflected in the correlations between the random intercepts. Secondly, this approach enables us to directly test our hypothesis that between-family differences in quantitative involvement moderate the interplay between qualitative involvement and emotional adjustment: Between-family differences in the interplay between qualitative parenting behaviors and emotional adjustment are estimated as random effects, and we can include between-family level moderators as predictors of these random effects. In this way, we included joint and relative quantitative involvement, and child sex as moderators. Figure \@ref(fig:figure1) shows a graphical representation of the resulting model. We conducted separate analyses for the four qualitative parenting dimensions, because a model including all four variables at once was too complex to estimate.

```{r figure1, out.width="100%", fig.cap="T-1: previous measurement occasion. ri: random intercept. s: random slope. M: Mother's qualitative involvement. E: Emotional adjustment. F: Father's qualitative involvement. inv: quantitative involvement. Thus, sME represents the random slope of Mothers' qualitative involvement on emotional adjustment."}
knitr::include_graphics("Figure1.png")
```

@@

tabchildeff$Parameter <- gsub("^\\((.+) on.*$", "\\1", tabchildeff$Parameter)
tabchildeff$Parameter[c(1,4)] <- c("Random intercepts of child effects on:", "Variance of child effects on:")
tab <- rbind(tabparenteff, tabchildeff)
#tab[is.na(tab)] <- ""
rownames(tab) <- NULL
reg_table(tab, caption = "Between-family level intercepts and variances of within-family predictive effects", note = "Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

We examined regression coefficients from joint and relative quantitative involvement to the random intercepts of qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). With regard to parenting, we found partial support for our hypotheses: For both parents, joint involvement positively predicted greater warmth, consistency, and inductive reasoning. Notably, the association between joint involvement and warmth was almost double in size for fathers, as compared to mothers. We did not find the hypothesized negative effects of joint involvement on parental anger. Counter to our hypothesis, we found that greater joint involvement was negatively associated with emotional adjustment. Thus, the children in families with highly involved parents generally displayed poorer emotional adjustment.

@@

```{r tabinvolvement, results = "asis"}
tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
tabinv$Parameter[c(1,5)] <- c("Effects of joint quantitative involvement on random intercepts of:",
                              "Effects of relative quantitative involvement on random intercepts of:")
tabinv$Parameter <- gsub("^(.+) on (?!random).*$", "\\1", tabinv$Parameter, perl= TRUE)
#reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
tabinvmod <- rbind(
  tab_std_sections$`Moderating effect of quantitative involvement on parenting effects`[c(1:3, 1, 4:5),],
  tab_std_sections$`Moderating effect of quantitative involvement on child effects`[c(1:3, 1, 4:5),])
tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on parenting effects of:",
                              "Moderating effects of relative involvement on parenting effects of:",
                              "Moderating effects of joint involvement on child effects on:",
                              "Moderating effects of relative involvement on child effects on:")
tab <- rbind(tabinv, tabinvmod)
#tab[is.na(tab)] <- ""
rownames(tab) <- NULL
cat(as.character(reg_table(tab, caption = "Standardized regression coefficients of between-family differences in quantitative involvement on between-family differences in qualitative parenting and emotional adjustment, and on within-family predictive effects", note = "The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")))
```

@@

To examine interactions between involvement quality and quantity, we tested whether between-family differences in quantitative involvement predicted the within-family cross-lagged effects between qualitative parenting behaviors and emotional adjustment (Table \@ref(tab:tabinvolvement)). With regard to the moderation of parenting effects, the only significant effects we found were for maternal warmth. Specifically, the within-families predictive effect of maternal warmth on emotional adjustment was negatively attenuated when both parents' joint quantitative involvement was higher. This same effect was positively attenuated, however, if mothers' involvement was relative higher than fathers' involvement. We found no moderation for paternal warmth, nor for both parents' consistency, inductive reasoning, and anger.

With regard to the moderation of child effects by quantitative involvement, we found that the effect of adjustment on maternal warmth was negatively attenuated when parents' joint quantitative involvement was higher, and also when mothers' relative involvement was greater compared to fathers. We found no effects for consistency. For inductive reasoning, we found that that the cross-lagged effect of adjustment on maternal inductive reasoning was negatively attenuated when joint involvement was higher. Adjustment was also more negatively predictive of fathers' inductive reasoning when mothers were relatively more involved than fathers. Thus, in families where mothers were more involved than fathers, children's increased emotional adjustment more strongly predicted decreased paternal reasoning - and conversely, decreased emotional adjustment predicted increased paternal reasoning. We found no moderating effects for anger.

@@

```{r tabsex, results = "asis"}
tabsex <- rbind(
  tab_std_sections$`Sex differences in parenting and adjustment`,
  tab_std_sections$`Sex differences in cross-lagged effects`[c(1:3, 1, 4:5),])
tabsex$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabsex$Parameter)
tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
tabsex$Parameter[c(1,5,8)] <- c("Sex differences in parenting and adjustment",
                              "Sex differences in parenting effects of:",
                              "Sex differences in parenting effects on:")
#tabsex[is.na(tabsex)] <- ""
rownames(tabsex) <- NULL
reg_table(tabsex, caption = "Standardized regression coefficients of child sex on average levels of qualitative parenting and emotional adjustment, and on within-family predictive effects.", note = "Sex is a binary predictor (-1 = female, 1 = male). The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

@@

Our operationalization of quantitative involvement, in terms of joint involvement and relative involvement, allowed us to addresses the claim that any effects of father involvement on children's adjustment might in fact be due to children receiving greater total care [@deutschPaternalParticipationChild2001]. However, counter to our hypothesis, joint involvement was *negatively* associated with between-family differences in emotional adjustment. One potential explanation is suggested by self-determination theory: Excessive levels of involvement might undermine children's autonomy needs, thereby leading to increased emotional maladjustment [@vandergiessenPerceivedAutonomySupport2014]. Research on "helicopter parenting" supports this interpretation, as psychological well-being was found to be lower in children of overly involved parents [@schiffrinHelpingHoveringEffects2014]. Another potential explanation is that the amount of quantitative care provided is not only driven by parents' motivation and ability to be involved in their children's lives, but also by children's objective needs. Factors related to children's objective needs - such as developmental delay, or atypical development - might thus drive an increase in parents' quantitative involvement, and might also be associated with children's emotional maladjustment, thereby resulting in a negative correlation between the two.

@@

We found limited evidence that quantitative involvement moderated within-family parenting effects. One such effect is that the effect of maternal warmth on adjustment was contingent on both joint- and relative quantitative involvement. Mothers' warmth predicted emotional adjustment less strongly when both parents were more involved, and more strongly when mothers were relatively more involved than fathers. These findings suggests a compensatory effect, whereby mothers' warmth can make up for disadvantages incurred by having a relatively uninvolved father [@shinnFatherAbsenceChildren1978].

The evidence that quantitative moderates child effects was somewhat stronger. Firstly, when joint involvement was higher, and when mothers' relative involvement was greater compared to fathers, emotional adjustment became less predictive of maternal warmth. These findings can be interpreted in the context of the sizable average positive effect of adjustment on maternal warmth. In families where mothers' involvement was high - either in terms of joint, or relative involvement - this positive effect was attenuated somewhat. Secondly, joint quantitative involvement positively attenuated the link from emotional adjustment to maternal inductive reasoning. This suggests that mothers have more of a role as disciplinarian in families where both parents are highly involved. When mothers were relatively more involved than fathers, fathers decreased their inductive reasoning more strongly in response to children's positive adjustment, and increased it more strongly in response to maladjustment. Together, these findings suggests that fathers take more of a role as disciplinarian when mothers are relatively more involved in care tasks.

@@

Nowadays, it is commonly acknowledged that parents and children reciprocally influence one another. However, the methods used to demonstrate bi-directional effects have been criticized for misrepresenting the direction of effects when stable between-family differences are not accounted for [@hamakerCritiqueCrosslaggedPanel2015]. We are aware of only one prior study that controlled for stable between-family differences when examining the interplay between parenting and emotional adjustment. This study found that child effects substantially outnumbered parenting effects in adolescence [@vanlissaRoleFathersMothers2018a]. The authors explained this finding in light of the relative equality of parent-adolescent relationships. The present study casts doubt on that interpretation, as we found a preponderance of child effects between the ages of 6-12. Although it is, of course, possible that parent effects are in effect at an even younger age, we can also consider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?

The finding that parent effects become smaller or non-significant once stable between-family differences are accounted for has also been demonstrated in other areas of developmental research (REF). As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory. Such findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time. Methodological innovations that examine the role of stable between-family differences can thus inspire more sophisticated theoretical conceptualizations of parental influences in child development.

At the same time, we should consider the limitations of these methods. The assumption that within-family fluctuations adequately capture all causal influences that parents exert on their children, for instance, is questionable. Stable factors can also exert changing influences over time. As an illustrative example: The genetic makeup that is passed from parents to children is stable, and yet the expression of some genes emerges over time. It might similarly be possible for stable parental psychosocial characteristics to influence children differently at different time points. Another important limitation is that within-family cross-lagged effects are time scale dependent. If parents shape their children through specific behaviors in dyadic interactions, then annual reflective self-reports might fail to capture parenting effects. Thus, although these methodological innovations are undoubtedly a step forward, they might not be the definitive paradigm for investigating the interplay between parents and children.

@@

Several strengths allow us to make unique contributions to the literature. First, mothers and fathers both completed identical parenting measures, which allowed us to disentangle and compare their unique roles in children's emotional adjustment development. Second, we incorporated both quantitative and qualitative involvement. This allowed us to provide a more nuanced overview of the links between parental involvement and emotional adjustment development than prior work, including how the distribution of quantitative childcare relates to parenting quality and emotional adjustment, and moderates links between these two constructs. Although some prior studies have also incorporated both quality and quantity of parenting, this work did not disentangle the total amount of care children received from the relative distribution of care tasks [see @deutschPaternalParticipationChild2001]. We addressed this limitation by estimating the latent mean and difference of absolute maternal and paternal quantitative involvement. This allows us to disentangle the effect of joint involvement, which reflects the total amount of care children receive, from relative involvement, which reflects the childcare distribution. Third, we used innovative methods to examine the extent to which parenting and adjustment are linked at the between- versus within-family level. Compared to methods that do not distinguish between these levels, our approach estimates dynamic reciprocal processes within the family more accurately. Our findings contribute to a growing literature suggesting that stable between-family differences might play a larger role than was previously thought, even in early childhood. Confidence in the results is bolstered by the fact that we used multi-informant reports: Each parent reported on their own parenting, and teachers reported on children's emotional adjustment. This avoids common method bias and single reporter bias. As teachers are likely to be "blind" to parents' child-rearing behavior, and have no stake in the parent-child relationship, their reports might be less biased in general.

@@

<!--  -->

@@

We found substantial evidence that both parents reported engaging more in behaviors that are positively associated with children's adjustment when joint involvement was higher. Conversely, when mothers were more involved than fathers, mothers' expressed more warmth, but fathers displayed less desirable parenting behaviors (less consistency, inductive reasoning, and warmth). Children's emotional adjustment, too, was lower when mothers were more involved than fathers. These findings highlight the important of considering interactions between quantitative and qualitative involvement, and suggest that all family members might benefit from a more equal distribution of childcare tasks.
